Distinguishing Between Spin-Aligned and Isotropic Binary Black Hole Populations Using Gravitational Wave Observations

### 2017-03-27 paper/AlignedIsotropicGW.tex
Changed "increasing" to "high", "decreasing" to "low" to avoid letter-collision.
Also implemented distribution plot comparing q = 1 and q = 0.5 mass ratios.  No appreciable change.

\newcommand{\OOneSigmaIsoAligned}{\checkme{1.7}}
\newcommand{\OOneOddsIsoAligned}{\checkme{0.087}}
\newcommand{\OTwoSigmaIsoAlignedMin}{\checkme{3.0}}
\newcommand{\OTwoOddsIsoAlignedMin}{\checkme{0.0012}}

@@

Each of our models for the merging \ac{BBH} spin population assumes
that the merging black holes are of equal mass (this is marginally
consistent with the three observations \citep{O1-BBH}, and the
$\chieff$ distribution is not particularly sensitive to the mass ratio
between the merging objects---see Section \ref{sec:mass-ratio}).  We
assume that the population spin distribution factorises into a
distribution for the spin magnitude, $a$, and a distribution for the
spin angles.  Finally, we assume that the distribution of spins is
common to each component in a merging binary.  Choosing one of three
magnitude distributions
\begin{equation}
  \label{eq:magnitude-dists}
  p(a) = \begin{cases}
    2\left(1-a\right) & \textnormal{``decreasing''} \\
    1 & \textnormal{``flat''} \\
    2\left(a-1\right) & \textnormal{``increasing''}
  \end{cases},
\end{equation}
and pairing with either an isotropic angular distribution or a
distribution that generates perfect alignment with the positive $z$
axis yields six different models for the $\chieff$ distribution.
These models are shown in Figure \ref{fig:chieff-distribution-models}.

\begin{figure}
  \plotone{../plots/chi-eff-distributions}
  \caption{\label{fig:chieff-distribution-models} The models for the
    distribution of $\chieff$ considered in this paper.  In all models
    we assume that the binary mass ratio $q \equiv m_1/m_2 = 1$ and
    that the same distribution of spin vectors obtains for each
    component.  The ``flat'' (blue lines), ``increasing,'' (green
    lines), and ``decreasing'' (red lines) magnitude distributions are
    defined in Eq.\ \eqref{eq:magnitude-dists}.  Solid lines give the
    $\chieff$ distribution under the assumption that the orientations
    of the spins are isotropic; dashed lines give the distribution
    under the assumption that both objects' spins are aligned with the
    orbital angular momentum.  The isotropic distributions are readily
    distinguished from the aligned by the production of negative
    $\chieff$ values, while the distinction between the three models
    for the spin magnitude distribution is less sharp.}
\end{figure}

@@

\begin{figure}
  \plotone{../plots/Wills_evidence_ratio_figure_with_mixture_models.png}
  \caption{Odds ratios among our models using the approximations to
    the posteriors on $\chieff$ from the O1 observations shown in
    Figure \ref{fig:O1-posteriors}.  The flat (``F''), increasing
    (``I''), and decreasing (``D'') spin magnitude distributions (see
    Eq.\ \eqref{eq:magnitude-dists}) are paired with isotropic,
    aligned angular distributions (``$\mathrm{I}$'',
    ``$\mathrm{A}$''), as well as a mixture model of the two
    (``$\mathrm{M}$'').  The most-favoured models have a decreasing
    distribution of spin magnitudes.  The odds ratio between these
    models is $\OOneOddsIsoAligned$, or $\OOneSigmaIsoAligned\sigma$.
    For all magnitude distributions the pure-isotropic models are
    preferred over the mixture models; correspondingly, the posterior
    on the mixture fraction peaks at 100\% isotropic.}
  \label{fig:O1-odds}
\end{figure}

@@

\begin{figure}
  \plotone{../plots/six-way-O2-model-selection}
  \caption{\label{fig:O2-predictions} Distribution of odds ratios
    predicted with 10 additional observations above the three
    discussed in Section \ref{sec:O1}.  Each panel corresponds to
    additional observations drawn from one of the $\chieff$
    distribution models.  The model from which the additional
    observations are drawn is outlined in red.  The height of the blue
    bar gives the median odds ratio relative to the model from which
    the additional observations are drawn; the green line gives the
    68\% ($1 \sigma$) symmetric interval of odds ratios over 1000
    separate draws from the model distribution.  The closest ratio
    between the most-favoured isotropic model and the most-favoured
    aligned model is $\OTwoOddsIsoAlignedMin$, corresponding to
    $\OTwoSigmaIsoAlignedMin\sigma$ preference for the correct angular
    distribution; most models result in more than $5\sigma$ preference
    for the correct angular distribution.  Because the three
    observations from Section \ref{sec:O1} are included in each data
    set the ``correct'' model is not necessarily preferred over the
    others, particularly when that model uses the ``increasing''
    magnitude distribution, which is strongly dis-favoured from the O1
    observations alone.}
\end{figure}

>>>>

\newcommand{\OOneSigmaIsoAligned}{\checkme{1.8}}
\newcommand{\OOneOddsIsoAligned}{\checkme{0.08}}
\newcommand{\OTwoSigmaIsoAlignedMin}{\checkme{3.0}}
\newcommand{\OTwoOddsIsoAlignedMin}{\checkme{0.0012}}

@@

Each of our models for the merging \ac{BBH} spin population assumes
that the merging black holes are of equal mass (this is marginally
consistent with the three observations \citep{O1-BBH}, and the
$\chieff$ distribution is not particularly sensitive to the mass ratio
between the merging objects---see Section \ref{sec:mass-ratio}).  We
assume that the population spin distribution factorises into a
distribution for the spin magnitude, $a$, and a distribution for the
spin angles.  Finally, we assume that the distribution of spins is
common to each component in a merging binary.  Choosing one of three
magnitude distributions
\begin{equation}
  \label{eq:magnitude-dists}
  p(a) = \begin{cases}
    2\left(1-a\right) & \textnormal{``low''} \\
    1 & \textnormal{``flat''} \\
    2\left(a-1\right) & \textnormal{``high''}
  \end{cases},
\end{equation}
and pairing with either an isotropic angular distribution or a
distribution that generates perfect alignment with the positive $z$
axis yields six different models for the $\chieff$ distribution.
These models are shown in Figure \ref{fig:chieff-distribution-models}.

\begin{figure}
  \plotone{../plots/chi-eff-distributions}
  \caption{\label{fig:chieff-distribution-models} The models for the
    distribution of $\chieff$ considered in this paper.  In all models
    we assume that the binary mass ratio $q \equiv m_1/m_2 = 1$ and
    that the same distribution of spin vectors obtains for each
    component.  The ``flat'' (blue lines), ``high,'' (green
    lines), and ``low'' (red lines) magnitude distributions are
    defined in Eq.\ \eqref{eq:magnitude-dists}.  Solid lines give the
    $\chieff$ distribution under the assumption that the orientations
    of the spins are isotropic; dashed lines give the distribution
    under the assumption that both objects' spins are aligned with the
    orbital angular momentum.  The isotropic distributions are readily
    distinguished from the aligned by the production of negative
    $\chieff$ values, while the distinction between the three models
    for the spin magnitude distribution is less sharp.}
\end{figure}

@@

\begin{figure}
  \plotone{../plots/Wills_evidence_ratio_figure_with_mixture_models.png}
  \caption{Odds ratios among our models using the approximations to
    the posteriors on $\chieff$ from the O1 observations shown in
    Figure \ref{fig:O1-posteriors}.  The flat (``F''), high (``H''),
    and low (``L'') spin magnitude distributions (see Eq.\
    \eqref{eq:magnitude-dists}) are paired with isotropic, aligned
    angular distributions (``$\mathrm{I}$'', ``$\mathrm{A}$''), as
    well as a mixture model of the two (``$\mathrm{M}$'').  The
    most-favoured models have a decreasing distribution of spin
    magnitudes.  The odds ratio between these models is
    $\OOneOddsIsoAligned$, or $\OOneSigmaIsoAligned\sigma$.  For all
    magnitude distributions the pure-isotropic models are preferred
    over the mixture models; correspondingly, the posterior on the
    mixture fraction peaks at 100\% isotropic.}
  \label{fig:O1-odds}
\end{figure}

@@

\begin{figure}
  \plotone{../plots/six-way-O2-model-selection}
  \caption{\label{fig:O2-predictions} Distribution of odds ratios
    predicted with 10 additional observations above the three
    discussed in Section \ref{sec:O1}.  Each panel corresponds to
    additional observations drawn from one of the $\chieff$
    distribution models.  The model from which the additional
    observations are drawn is outlined in red.  The height of the blue
    bar gives the median odds ratio relative to the model from which
    the additional observations are drawn; the green line gives the
    68\% ($1 \sigma$) symmetric interval of odds ratios over 1000
    separate draws from the model distribution.  The closest ratio
    between the most-favoured isotropic model and the most-favoured
    aligned model is $\OTwoOddsIsoAlignedMin$, corresponding to
    $\OTwoSigmaIsoAlignedMin\sigma$ preference for the correct angular
    distribution; most models result in more than $5\sigma$ preference
    for the correct angular distribution.  Because the three
    observations from Section \ref{sec:O1} are included in each data
    set the ``correct'' model is not necessarily preferred over the
    others, particularly when that model uses the ``high'' magnitude
    distribution, which is strongly dis-favoured from the O1
    observations alone.}
\end{figure}
